### 2013-05-31 front-matter.tex
minor fix in abstract

\title{
    AA: the Algorithmic Auto-regulation (Software Development) Methodology
}

\author{%
Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negr\~{a}o \and Lucas Zambianchi
\and Marcos Mendonca \and Danilo Shiga
}

\maketitle
%\thispagestyle{empty}

\begin{abstract}
We present a new methodology for coordinating teamwork called Algorithmic
Auto-regulation (AA). This methodology is based on recent social networking
concepts and individual merit and exhibits convenient asynchronous,
auto-documenting and auto-regulating characteristics. Team members take on an
egalitarian and voluntary role, by logging into periodic ``AA sessions'' for an
arbitrary duration (e.g.\ 2 hours per day).  During each session, a user creates
a log composed of short text sentences about their activity. These logs are
publicly aggregated in a Website and are peer-validated, as in code review. A
short screen-cast is ideally recorded at the end of each session to make AA logs
more understandable. This methodology seems to be well-suited for increasing the
efficiency of teams working on Global Software Development (GSD). Recognized
reasons for this are the: 1) built-in asynchronous on-demand communication,
documentation and proven working hours; 2) reduced need for central management,
meetings and time-consuming reports. Hence, AA legitimizes and eases activities
of a distributed software team.  It enables groups to have new means to fund
activities, allowing for business models to emerge from distributed software
development. AA is proposed, at it's core, as a way of having self-replicating
hacker initiatives. These claims are discussed in a real case-study of a
distributed free software hacker team called Lab Macambira.
\end{abstract}