### 2013-06-01 paper.tex
revisado artigo todo na escrita

%\documentclass[letterpaper,peerreview,12pt,compsoc,draftcls]{IEEEtran}
%\documentclass[10pt,twocolumn,letterpaper]{article}
\documentclass[letterpaper]{article}

%
% Final Checklist TODO XXX 
% - Spellcheck
% - check xxx's todo's etc
%
%\input{preamble}
\usepackage{url}
\usepackage{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{color}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% You have two versions of the macro
% \draftnote{My note}. The first version puts notes (e.g. My note in the example)
% into the margin of your document. The second formats the note as nothing. You
% 'comment out' the version of the macro you don't want (using a % at the
% beginning of the line).
%\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
%\newcommand{\draftnote}[1]{}

% This one is just for the comments for in-line text.
\newcommand{\indraftnote}[1]{\textcolor{blue}{\texttt{\footnotesize[#1]}}}
\newcommand{\todo}[1]{\indraftnote{todo: #1}}
%\newcommand{\indraftnote}[1]{}

\begin{document}
\input{front-matter}

\section{Introduction}

One of the defining features of modern times is the widening geographical
distribution of software teams~\cite{last2003}. This is responsible for the so
called Global Software Development (GSD)~\cite{german2003}.  An example is the
free software movement. Projects and institutions like Mozilla Foundation has
several employees and thousands of voluntary developers distributed across many
countries. The same is true for GNOME~\cite{german2003}, OpenBSD, MySQL or
Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source
  network, have a more complete and constantly updated list of the most active
  projects on-line at \url{www.ohloh.net}}. Along the free and open software
projects, GSD has a growing popularity in every niche of the software industry
as a whole, even on those distributing their software with proprietary
licenses. This phenomenon is attributed to a variety of factors such as a larger
labor pool, natural globalization of software companies and foundations or even
the premise of cheaper cost of production~\cite{komi2005}.

Despite the advantages of GSD, it is known how difficult it is to coordinate and
fund free software on a larger scale than currently available. That is, when
teams are very heterogeneous, containing not only volunteers and experienced
developers, but also contractors from different backgrounds and cultures. This
observations are in consonance with dedicated studies, such
as~\cite{carmel1999}.  As a common base, related difficulties for GSD gather
around distance, time and cultural differences. In the case of free or open
software projects, all these factors are involved.

Another problem faced by modern software companies and other collectives are
frequent ineffective meetings, which are seldom focused on the interest of any
attendant. The result is that it has become the norm to participate in these
with the ``laptop open'', which is unproductive. Software developers like to
code, to be productive, to have their hands on their project, to do what they
are best at. They dislike to have to stop for meetings or to write lengthy
reports to justify their funding.
%\todo{ler mais. cacm, etc}

To address these matters is the purpose of AA, a methodology and an associated
software system for coordinating distributed team work dealing with the
disadvantages of GSD. Team members take on an egalitarian role, and stay
voluntarily \textit{logged} in the system for part of their time (e.g.\ 2 hours
per day), during which they log a periodical short text sentence --- similar to
a ``tweet'' in this Twitter era --- as the status of their activity using an
easy to use command. These ``micro-blog sentences'' are publicly aggregated and
validated by other team members. Through AA, the community has a methodology and
an associated system to validate and enable the activities of a distributed
software team. It implicitly legitimizes financial support for the expansion of
the activity of the developer team as the activities are documented and
validated.  The AA methodology is specially useful for coordinating distributed
and decentralized team work, providing effective means to asynchronously update
different team members, diminishing the need for harsh synchronous meetings.

%\todo{alerts work for greater consciousness of time}
%\todo{Integrate notes from Ricardo+Renato april 15/16 meeting}
%\todo{integrate IRC chat notes}

A brief overview of current papers about GSD methodologies related with AA is
presented in section~\ref{related-work}. In section~\ref{aa-methodology}, the
most relevant characteristics of the AA methodology are outlined. In this same
section, there is a description of an AA use in a team of 9 developers in the
second half of 2011 and a broader a more released use of AA from 2012 until
now. Finally, in Section~\ref{conclusions}, there are final conclusions and
indicatives of future possibilities to the practical use of AA on other teams of
software developers or individuals working on non-software activities.

\section{Related work}
\label{related-work}

%% aqui tomei como ponto de vista as metodologias associadas a GSD
%% (Global Software Development) que considero foco do AA, e sua
%% principal vantagem

%\todo{survey other methodologies such as agile etc}
%\todo{probably a good source http://agilemanifesto.org/}

There has been a large amount of research in methodologies to
deal with distributed teams of developers. Although this article focuses in GSD,
some principles involved on those methodologies could be used on smaller teams
of developers working in the same place, time period and with minor cultural
differences. Moreover, ``distributed development'' is generally tough as being
global which is not true. AA was applyed to a team that work at the
same city but on different times (more details on Section~\ref{results}). Even
smaller groups of developers working on the same building could use GSD
methodologies. A thorough survey of these methodologies is beyond the scope of
this paper. This section presents a brief overview.

Various methodologies for GSD were built around the factors that affect
distributed team works. As proposed by Carmel~\cite{carmel1999}, these comprises
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team needs
to have good communication to work on tasks related to a specific problem. Cultural
distance encompass differences on organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). Finally, the temporal distance
prejudices synchronous communications like telephone or video-conferences.
Developers working on different time-zones manage
their agenda guided by this temporal distance.

Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
enhance intensive collaboration. His approach divides the whole software
life-cycle on levels of complexity. Each level has a degree of
collaboration. For example, some developers working on a project with high
collaboration level should use the follow-the-sun approach: when concluding the
work day, they pass their work to the team working on another time-zone. Other
tactics are suggested by the same author to deal with the three distances, like
separating foreign units of developers in time-zone bands.

Battin et al.~\cite{battin2001} argues about their experiments using
methodologies created for the development centers of the
Motorola Company (which has over 25 software development centers
worldwide). These methodologies included constant communication with critical
units, incremental integration and schedules based on time-zones distributed to
developers on 6 countries from 3 continents.

While considering free software projects instead of companies, similar factors
are present and some methodologies arises. German~\cite{german2003} gives a
concise review of methodologies used by the GNOME project, one of the most
active free software projects and used by companies like Sun Microsystems. It is
interesting to note a difference on the viewpoint present on his paper: German
focuses the methodology description on code. He start explaining that GNOME is
separated into modules (76 on version 2.4, to be precise) and each module has
one maintainer who divide his modules into separated parts in which other
developers can work on independent tasks, along other responsibilities. As like
others free software projects, all the development was done using a
\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
Chat (IRC) to discussion and communication and a software version
management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
is common on free and open source projects for face-to-face meetings and is based
each time on a different place.

\section{The AA methodology}
\label{aa-methodology}

As noted, some strategies for GSD is based on complex methodologies and many of
those were built for a specific company or software center. This section proposes an
alternative based on a simple idea: small working sessions logged by
a computational tool. Figure~\ref{fig:mm} summarizes this methodology.

\begin{figure}
\begin{center}
   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm2.png}
\end{center}
   \caption{
   Mindmap of AA methodology: 1) developer engagement cycle, i.e. his use of AA; 2) functionality as main goals of the system; 3) potentialities as enhancements AA delivers to the developer's medium. 
   }
\label{fig:mm}
\end{figure}

\subsection{The AA session}

From a developer's perspective, the AA methodology is based on creating
small, high perspective, reports of what is being done in a specific time frame. This frame
can be something between 5 to 15 minutes, depending on what is 
comfortable for the developer. An AA Session is a period of about 2
hours, in which these short reports (called ``shouts'') are periodically written. The developer can set reminders to show up when it is
time to make a new report. The objective of the flexible time frame and
reminders is to minimize developer overhead during his AA session. In this way,
he can make the reports with little or no loss of concentration on his coding activity. Each report can
be sent directly to an online server, or stored locally in a temporary
database for sending later on. This makes it possible to use AA while offline.

Developers can also record a video screencast at the end of the session,
summarizing what has been done, explaining with his words and showing on the screen his most
important results. This, combined with the textual log of his session, makes the
whole report more significant as documentation and understandable for himself or other people searching for
information about his production.

\subsection{The AA website report}

\begin{figure}
\begin{center}
   \includegraphics[width=0.95\linewidth]{figs/aa-0_1_.png}
\end{center}
   \caption{AA Version 0.1. Messages of users: hybrid, filter0, v1z e aut0mata about activities of interest for labMacambira.sf.net and other parties (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla, PulaPirataComics). Each message is a ``shout'', which, grouped, forms an AA session.}
\label{fig:aaserver}
\end{figure}

All AA reports made by the developers are sent to a Web server and become public
in a Website. It is then possible for a manager or another developer to follow
closely the work of a developer, nearly real-time, reading each of his
small reports of what he is doing.

Another possibility is to check older sessions to see what was done
and the comments of the developer about it. Since each AA post (or ``shout'') happens in a
small time frame of work, the information about what was done becomes easy
to understand, which seems better than a long report in the end of a session.

The developer can add to his session a link for his screencast
about that session. This screencast is a video summary of what was done and complements the report.

\subsection{Validation}
Each AA session must be validated by another developer. This means that all
reports are read by someone that will consider valid or not and will even
write comments about the specific session. The developer in charge of
validating a session is decided randomly by the AA Web server, which sends an
email to the chosen developer with the URL to a validation interface.

% TODO: adicionar figura aa cliente (ou tabela com os principais comandos)
% TODO: adicionar figura com arquitetura do AA cliente + AA web + validação
% TODO: adicionar figura do pAAinel

\section{Results and discussion}
\label{results}

The easy and effective management of teams working on GSD is the main purpose of
the AA methodology. This methodology was applied to a group of 9 developers in the second half of 2011 on what was named ``Lab Macambira''\footnote{LabMacambira.sf.net:
  \url{http://labmacambira.sf.net}}. The main objective of the team was to work
on different free software projects, contributing directly to their development,
sending bug corrections or proposing new features on their source code.

Team members earned knowledge on software development
as part of large and distributed free software projects like Scilab and  
Mozilla. In this way, one month of training was conducted by three experienced
developers, teaching the basics of development support tools like
bugtrackers, programming languages and version control systems. After this
period, a challenge was proposed for the new developers: send a bug correction
or a new feature to a large free software project and you will work with us, you
will be a ``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
the contributions of each ``Macambira'' to free software projects.

%% adicionar tabela de contribuições

\begin{table}
  \caption{Free and open software projects that received contributions
    from ``Macambiras''. On the first column there is a list of
    applications. At right, the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    ``Lab Macambira'', and at free software community in general, is a
    common practice to use a pseudonym as identification.}
    \begin{center}
    \small\begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
\end{tabular}
\end{center}
    \label{tabela:contribuicoes}
\end{table}

In one month, ``Macambira'' developers contributed to many large free software
projects. Some of the developers started the training with no knowledge of what
is free software and ended that period becoming a free software developer.

During that month, the same team developed the first version of the AA system
and used AA to manage their activities, even while developing the system. All
the source code of AA --- the client that sends the logs and the AA Web server
--- is publicly available\footnote{AA source code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the AA session logs of the ``Lab Macambira'' team is also
online\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the training period, during 6 more months, the ``Macambiras'' worked on a
large range of free software projects, distributed on work groups. Each work
group focused on a specific theme like video, audio and web. They were financed by
contracts and support of the ``Pont\~{a}o de Cultura Digital: N\'{o}s Digitais''. In
Table~\ref{tabela:criados} we can see a list of the free software created by
``Lab Macambira'' since July of 2011.

\begin{table}
    \caption{Software projects, created by ``Lab Macambira'', with the use of AA, since July
      of 2011 with a short description and the technologies --- like
      programming languages or frameworks --- involved. It is
      interesting to note the heterogeneity of projects and its areas
      of application.}
    \small\begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}
    \label{tabela:criados}
\end{table}

As of this writing the ``Lab Macambira'' have many software developers, and some
of the trained developers continue to work voluntarily in the project.

\section{Conclusions}
\label{conclusions}

%% brief introduction
In a scenario where GSD is growing as a popular form of software development,
not just on free software projects but in the whole software industry, there is an increased need for
methodologies to deal with it's disadvantages and at the same time to amplify it's
advantages.

This paper has presented a simple methodology to GSD, being the development conducted
on large or small groups of software developers, working on different countries
or even at the same room. The AA methodology implements a simple system where
each developer takes notes of his work, generating a periodical log of small text
sentences. The sum of those sentences, along an entire session of work, results
in a complete report. The report is made public available through a Website and
are validated by other developers sorted aleatory by the AA Web server.

Instead of a merely work-management tool, AA acts as a methodology to improve the
time sense of individuals, dividing their work on small sessions, and also
reducing the need of extensive reports or unnecessary meetings. By asking users
to write a minimal text sentence as a continuous log, AA minimizes
disturbances: developers just type some
characters, hit \textit{enter} and go back to coding, which can help mental organization about the work being done.

AA application is not restricted to software development. As of this writing
there is a comic book studio\footnote{Pula Pirata Comics: \url{http://pulapirata.com}}
using AA to manage their activities. People with non-software
background, like social scientists, musicians and activists has also been using AA
and contributing for its improvement.

For software developer teams, the use of AA was used successfully to auto-regulate the work
of ``Lab Macambira'', a group of free software developers from Brazil. Since
July of 2011 the group have contributed and created new free and open source
software for a vast number of applications.

There are many aspects of the work which remain unfinished. New ways to report
logs --- the ``Twitter like'' messages, sometimes called ``shouts'' --- from different interfaces like IRC,
Internet Messaging services and email can make the use of AA easy and
widespread, turning AA an ubiquitous system, presented on everyday communication
channels. Even the work logs generated since July of 2011 could be 
statistically analyzed to recognize patterns in the behavior of
individuals and their productions.

Being a free software
system and an open methodology, AA could be used to auto-manage groups
working on software or other kinds of activities. In this way,
there is interest on spreading AA to have even more developers
contributing in a collaborative way.

\section*{Acknowledgments}
%The authors would like to thank NSF and CNPq.

Authors thank AA: the present research and even this manuscript was written using AA. The complete log is on-line at
\url{http://www.pulapirata.com/skills/aa}.
Authors are also grateful to IFSC/USP, IPRJ/UERJ, IFT/UNESP and all AA users and collaborators, spacially those who coded AA hacks for shell and bot uses and those who coded the different Web interfaces in current use.

\nocite{last2003}
\nocite{german2003}
\nocite{carmel1999}
\nocite{carmel2001}
\nocite{komi2005}
\nocite{battin2001}

%{\small
%\input{paper-draft.bbl}
%%%%\bibliographystyle{splncs}
% this is a good style for drafting
%\bibliographystyle{abstract}
% this is a good style for finals
\bibliographystyle{acm}
\bibliography{aa.bib}
%}

\end{document}